---
title: "HW #4 "
output: html_notebook
---

#1. For two of these variables (that are not symmetric), find a transformation which makes the dataset symmetric.  You can use any tool you want, including H quick method, inspection of mids, symmetry plots, etc.  Demonstrate that your transformation has been symmetric.
```{r}
library(LearnEDAfunctions)
head(lake)
```
```{r}
aplpack::stem.leaf(lake$Area)
```
```{r}
lval(lake$Area)
```
The data is right-skewness
```{r}
symplot(lake$Area)
```
To remove the right-skewness that we see, we go down the ladder of powers and try powertransformationspthat are smaller thanp= 1.
```{r}
roots <- sqrt(lake$Area)
aplpack::stem.leaf(roots)
```
```{r}
symplot(roots)
```
We can see it also has some plots above the line.
Let’s continue down the ladder of powers and try thep= 0power (logs). Again, we show thestemplot, the letter-value display and the symmetry plot.
```{r}
logs <- log(lake$Area)
aplpack::stem.leaf(logs)
```
```{r}
symplot(logs)
```
It can be seen that compared to the above figure, the data is obviously closer to symmetry.
```{r}
aplpack::stem.leaf(lake$Depth)
```
```{r}
lval(lake$Depth)
```
```{r}
symplot(lake$Depth)
```
We can see that all the points are above the straight line, so the data is skewed to the right.To remove the right-skewness that we see, we go down the ladder of powers and try powertransformationspthat are smaller thanp= 1.
```{r}
roots <- sqrt(lake$Depth)
aplpack::stem.leaf(roots)
```
```{r}
symplot(roots)
```
We can see obvious changes in the symmetry graph, and some points begin to appear on the straight line. The data began to become symmetrical.
Let’s continue down the ladder of powers and try thep= 0power (logs). Again, we show thestemplot, the letter-value display and the symmetry plot.
```{r}
logs <- log(lake$Depth)
aplpack::stem.leaf(logs)
```
```{r}
symplot(logs)
```
We can see from the distribution of points that the distribution of data is more even and symmetric.
#2.Find two datasets (at least 30 observations in each) that you are interested in that are not symmetric.  Find a transformation that makes each dataset symmetric and demonstrate that your transformation is effective in achieving approximate symmetry.
```{r}
aplpack::stem.leaf(braves.attendance$Attendance)
```
```{r}
lval(braves.attendance$Attendance)
```
```{r}
symplot(braves.attendance$Attendance)
```
We can see that all the points are above the straight line, so the data is skewed to the right.To remove the right-skewness that we see, we go down the ladder of powers and try powertransformationspthat are smaller thanp= 1.
```{r}
roots <- sqrt(braves.attendance$Attendance)
symplot(roots)
```
Let’s continue down the ladder of powers and try thep= 0power (logs). Again, we show thestemplot, the letter-value display and the symmetry plot.
```{r}
logs <- log(braves.attendance$Attendance)
symplot(logs)
```
We can see from the distribution of points that the distribution of data is more even and symmetric.
```{r}
aplpack::stem.leaf(NationlNames2$Count)
```
```{r}
lval(NationlNames2$Count)
```
```{r}
symplot(NationlNames2$Count)
```
We can see that the data on the right side of the median has a right-skewed phenomenon.
```{r}
roots <- sqrt(NationlNames2$Count)
symplot(roots)
```
Let’s continue down the ladder of powers and try thep= 0power (logs). Again, we show thestemplot, the letter-value display and the symmetry plot.
```{r}
logs <- log(NationlNames2$Count)
symplot(logs)
```
We can see from the distribution of points that the distribution of data is more even and symmetric.
